---
title: "class1-d"
author: "Sujatha Jagannathan"
date: "February 6, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Preparing R
Load packages  
```{r, include = TRUE}
library("tidyverse")
```
This should install readr, tidyr, dplyr, purr, stringr, ggplot2, tibble, and forcats.

## 2. Importing data
1. RNA read counts  
```{r cars, echo = TRUE}
count_summary_rna <- read_delim("/Users/sjaganna/Desktop/CU-onedrive/OneDrive - The University of Colorado Denver/08-teaching/2018_MOLB7910/1-class1/data/count_summary_rna.tsv", 
                                delim = "\t", 
                                escape_double = FALSE, 
                                trim_ws = TRUE, 
                                skip = 1)
```

2. Experimental metadata
```{r}
experiment_metadata <- read_delim("~/Desktop/CU-onedrive/OneDrive - The University of Colorado Denver/08-teaching/molb7910/1-class1/data/experiment-metadata.txt",
"\t", escape_double = FALSE, trim_ws = TRUE)
```

## 3. Getting familiar with the data

```{r}
count_summary_rna #this works well with tbl_df data type. Others could take a LONG time  

# display first six rows of a dataframe
head(count_summary_rna) 

# display last six rows of a dataframe
tail(count_summary_rna)  
```

#### Str and Glimpse
```{r}
#Compactly display the internal structure of an R object
str(count_summary_rna)

#Get A Glimpse Of Your Data
glimpse(count_summary_rna)
glimpse(experiment_metadata)
```

#### Summary

A generic function used to produce result summaries of the results of various model fitting functions.
```{r}
summary(count_summary_rna)
summary(count_summary_rna$`0h_1_RNA_S13_umi.bam`)
```

#### Table
Uses the cross-classifying factors to build a contingency table of the counts at each combination of factor levels.
```{r}
table(experiment_metadata$hour)
```

## 4. Tidying data

- Rename columns
- Drop unnecessary columns
- Spread and gather

#### Renaming columns
```{r}
names(count_summary_rna)
```

#### Renaming columns (cont'd)
```{r}
# count_summary_rna <- count_summary_rna %>% rename(`0h_1_RNA_S13_umi.bam`, hour1_rep1)
```

## 5. Transform data

- Merge with gene_symbols