---
title: "Homework"
author: "Neelanjan Mukherjee"
date: "2/26/2019"
output: html_document
---
The goal of this homework is to perform exploratory data analysis on  DUX4 transcriptional targets.

Please execute the following tasks, knit the resulting .Rmd (Lastname_class2-homework.Rmd) and submit the .html output as your assignment.
You will need to import the following data:  
* `data_gene-level.tsv`. (Files: 2-class2 > data > input > data_gene-level.tsv)  
* `target_genes.csv` file provided (Files: 2-class2 > data > input > target_genes.csv)  

## __Perform the following for DUX4 targets only:__

1. Make boxplots and jitterplots of counts for all samples on the same plot. Please make sure to transform the y-axis (log10 or log2) for better visualization. Describe your result in a sentence or two.
2. Create heatmap of the correlation coefficient of between samples using `corrplot`. Please specify what metric you used and if you transformed the data. Describe your result in a sentence or two.
3. Create heatmap of clustered DUX4 target expression levels for all samples using `pheatmap`. Please specificy the `clustering_method` and `clustering_distance_rows`. Please make sure to __NOT__ to make a heatmap of the clustered correlation coefficients. Describe your result in a sentence or two.
4. Perform the same PCA analysis as in class except only for DUX4 targets. Describe your result in a sentence or two.
5. Always print `sessionInfo()`.
6. Knit and upload the resulting .html as your assignment. 






```{r setup}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)

# load necessary libraries
# setwd/here()

```

## Import data


```{r import}

# import data

```

## Box- and Jitter- plots

Make plots of counts accross all samples __only of DUX4 genes__. 

```{r plots}

# make plots

```

Super cool - the counts go up over time!!

## Correlation

Blah

```{r correlation}

# corrplot

```

Blah Blah

## Clustering DUX4 targets heatmap

Blah

```{r cluster}

# pheatmap

```

Blah Blah

## PCA DUX4 targets heatmap

Blah

```{r pca}

# pca

```

Blah Blah



```{r session info}

# print r session information

```
