---
title: "class3-homework.Rmd"
author: "Sujatha Jagannathan"
date: "February 28, 2019"
output: html_document
---
The goal of this homework is to get comfortable using and combining dplyr verbs to transform data. 

Please execute the following tasks, knit the resulting .Rmd (Lastname_class3-homework.Rmd) and submit the .html output as your assignment.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE, eval = TRUE)
```

#### 1. Load packages 
Packages: `tidyverse`, `skimr`, `here`, and `ggplot2` 
```{r, include = TRUE}
#load packages here

```

#### 2. Set working directory
```{r}
#setwd/here

```

#### 3. Import data
Path: "3-class3/data/input/data_transcript.csv""
```{r}
#import data

```

#### 4. select

1. Select the columns `hgnc_symbol`, `hour00_rep1`, and `hour14_rep1`; `glimpse` the resulting table.

```{r}
#select1

```

2. Select all columns except `Length`; `glimpse` the resulting table.

```{r}
#select2
```

#### 5. filter

1. Filter rows based on length of transcript > 1000; `glimpse` the resulting table. 

```{r}
#filter1

```

2. Filter rows based on hour14_rep1 > 5000; `glimpse` the resulting table.

```{r}
#filter2

```

#### 5. select & filter

`select` all columns except for `X1` and then pipe the output to `filter` to only retain rows where the `sum` of counts from the three hour00 replicates is > 20. `glimpse` the resulting table. 

```{r}
#select & filter

```

#### 6. mutate

Add a pseudocount of 10 to the counts from all time points and replicates. `glimpse` the resulting table.

```{r}
#mutate

```

#### 7. select, filter, and mutate

Select all columns except `X1`, filter rows based on hour14_rep1 > 5000; mutate by taking log10 of all count data (4 time points, 3 replicates). `glimpse` the resulting table. 

```{r}
#select, filter, and mutate

```

#### 8. group_by and summarise

Group rows by `hgnc_symbol` and summarise each of the count data (4 time points and 3 replicates) via `sum`. `glimpse` the resulting table. 

```{r}
#group_by and summarise

```

#### 9. select, group_by, summarise, mutate and filter

Select all columns except `X1`; Group rows by `hgnc_symbol` and summarise each of the count data (4 time points and 3 replicates) via `sum`; mutate by taking log10 of all count sums (4 time points, 3 replicates); `glimpse` the resulting table. 

```{r}
#select, group_by, summarise, mutate and filter

```

#### 10. Print sessionInfo {.smallest}

```{r}
#sessionInfo

```
