---
title: "class3-homework.Rmd"
author: "Sujatha Jagannathan"
date: "February 28, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE, eval = TRUE)
```

#### 1. Load packages 
Packages: `tidyverse`, `skimr`, `here`, and `ggplot2` 
```{r, include = TRUE}


```

#### 2. Set working directory
```{r}


```

#### 3. Import data
Path: "3-class3/data/input/data_transcript.csv""
```{r}


```

#### 4. select

1. Select the columns `hgnc_symbol`, `hour00_rep1`, and `hour14_rep1`; `glimpse` the resulting table.

```{r}


```

2. Select all columns except `Length`; `glimpse` the resulting table.

```{r}

```

#### 5. filter

1. Filter rows based on length of transcript > 1000; `glimpse` the resulting table. 

```{r}


```

2. Filter rows based on hour14_rep1 > 5000; `glimpse` the resulting table.

```{r}


```

#### 5. select & filter

`select` all columns except for `X1` and then pipe the output to `filter` to only retain rows where the `sum` of counts from the three hour00 replicates is > 20. `glimpse` the resulting table. 

```{r}

```

#### 6. mutate

Add a pseudocount of 10 to the counts from all time points and replicates. `glimpse` the resulting table.

```{r}

```

#### 7. select, filter, and mutate

Select all columns except `X1`, filter rows based on hour14_rep1 > 5000; mutate by taking log10 of all count data (4 time points, 3 replicates). 

```{r}

```

#### 8. group_by and summarise

Group rows by `hgnc_symbol` and summarise each of the count data (4 time points and 3 replicates) via `sum`. 

```{r}

```

#### 9. select, group_by, summarise, mutate and filter

Select all columns except `X1`; Group rows by `hgnc_symbol` and summarise each of the count data (4 time points and 3 replicates) via `sum`; mutate by taking log10 of all count sums (4 time points, 3 replicates); 

```{r}

```

#### 10. Print sessionInfo {.smallest}

```{r}

```
