---
title: "class3-presentation"
author: "Sujatha Jagannathan"
date: "February 6, 2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Learning goals

- Review material from class 1
- Understand the syntax of dplyr and tidyr
- Get comfortable with cleaning and transforming data
- Use in-class time to work through issues and answer questions

## Load packages  
```{r, include = TRUE, warning=FALSE, message=FALSE, tidy.opts=list(width.cutoff=60)}
library("tidyverse")
library("skimr")
library("here")
library("ggplot2")
```

## Set working directory
```{r, tidy.opts=list(width.cutoff=60)}
here() # prints out the path that `here` points to
setwd(here()) # set working directory to `here`  
getwd() # get working directory
```

## Import data
```{r}
data_transcript <- read_csv("3-class3/data/input/data_transcript.csv")

data_transcript <- read_csv(here("3-class3", "data", "input", "data_transcript.csv"))

data_transcript <- read_csv("/Users/sjaganna/Desktop/CU-onedrive/OneDrive - The University of Colorado Denver/08-teaching/molb7910/3-class3/data/input/data_transcript.csv")
```

# dplyr

## dplyr verbs

- `select`
- `rename`
- `filter`
- `arrange`
- `count`
- `mutate`
- `group_by` and `summarise`
- pipe: `%>%` 
- ... and many others! `add_row`, `add_column`, `distinct`, `sample_n`, `sample_frac`, `slice`, `pull`

## `select` and `rename`

- When do you use it?
- `select` to extract one or more columns from a table 
- `rename` to rename columns/variables

## `select`

- `select()` to extract one or more column from a table 

- Syntax: 
```{}
select(data = ..., <colname>)  
```
- Examples:   
```{}
select(data = data_transcript, hgnc_symbol)  
select(data_transcript, -X1)
select(data_transcript, starts_with("hour"))
select(data_transcript, hour00_rep1:hour14_rep3)
select(data_transcript, contains("ensembl"))
select(data_transcript, ends_with("rep2"))
select(data_transcript, dux4_target, everything())
```

## `select_if`

- `select_if()` to extract all columns of a particular type

- Example:
```{}
select_if(data_transcript, is.numeric)
select_if(data_transcript, is.logical)
select_if(data_transcript, is.character)
```

## `rename`

- `rename()` renames the variables in a table. Keeps all columns. `select()` can also do this, but with some caveats. 

- Syntax
```{}
rename(data = ..., <colname>)
```

- Examples:
```{}
rename(data_transcript, rownum = X1)
select(data_transcript, rownum = X1)
select(data_transcript, rownum = X1, everything())
```

## `filter`, `arrange`, and `count`

- When do you use it?
- `filter` to extract rows that meet a logical criteria. 
- `arrange` to order rows by values of one or more columns. 
- `count` to count the number of rows in which a value occurs within a column

## `filter`

- `filter()` chooses rows/cases where conditions are true.

- Syntax
```{}
filter(data = ..., <condition>)
```

- Examples:
```{}
filter(data_transcript, hgnc_symbol == "RPL28")
filter(data_transcript, hgnc_symbol %in% c("RPL28", "RPS3"))
filter(data_transcript, hour14_rep1 > 1000)
filter(data_transcript, hour14_rep1 > 1000 & hour14_rep2 > 1000)
filter(data_transcript, hour14_rep1 > 1000 | hour14_rep2 > 1000)
```

## `arrange`

- `arrange()` to order rows by values of a column or columns (low to high). `desc()` orders high to low. 

- Syntax
```{}
arrange(data = ..., <colname>)
```

- Examples:
```{}
arrange(data_transcript, dux4_target)
arrange(data_transcript, hour14_rep1)
arrange(data_transcript, dux4_target, hour14_rep1)
arrange(data_transcript, dux4_target, desc(hour14_rep1))
glimpse(arrange(data_transcript, dux4_target, desc(hour14_rep1)))
```

## `count`

- `count()` counts number of rows in each group defined by the specified variable.  

- Syntax
```{}
count(data = ..., <colname>)
```

- Examples:
```{}
count(data_transcript, dux4_target)
count(data_transcript, hgnc_symbol)
count(data_transcript, Length)
```

## `mutate`

- When do you use it?
- Use `mutate` to compute new columns and to transform existing columns. 

## `mutate` {.smaller}

- `mutate()` to compute new columns

- Syntax
```{}
mutate(data = ..., <newcolname> =  funs(<oldcolname>))
mutate(data = ..., <colname>, funs(x))
```

- Examples:
```{}
mutate(data_transcript, hour14 = (hour14_rep1 + hour14_rep2 + hour14_rep3)/3)
glimpse(mutate(data_transcript, hour14 = (hour14_rep1 + hour14_rep2 + hour14_rep3)/3))
mutate_at(data_transcript, vars(hour00_rep3), funs(log(.)))
mutate_if(data_transcript, is.numeric, funs(log(.)))
mutate_at(data_transcript, vars(starts_with("hour")), funs(log(.)))
```
Many other vector functions can be used with `mutate()` to compute new columns. Refer to cheatsheet. 

## `group_by` and `summarise`

- When do you use it?
- Use `group_by` to create a grouped copy of the table and `summarise` to then manipulate values within each group separately and then combine the results. 

## `group_by`

- `group_by()` to create a grouped copy of a table. 

- Syntax
```{}
group_by(data = ..., <colname>)
```

- Examples:
```{}
group_by(data_transcript, dux4_target)
group_by(data_transcript, hgnc_symbol)
```

## `summarise`

- `summarise()`  to create a new summary statistic

- Syntax
```{}
summarise(data = ..., <colname>)
```

- Examples:
```{}
summarise(data_transcript, mean_length = mean(Length))
summarise(data_transcript, hour00_rep1 = mean(hour00_rep1))
summarise_if(data_transcript, is.numeric, funs(mean(.)))
```
Many other summary functions can be used with `summarise()` to compute new columns. Refer to cheatsheet. 

## `group_by` + `summarise`  {.smaller}

- Syntax
```{}
group_by(data = ..., <colname>) %>% summarise()
```

- Examples:
```{}
group_by(data_transcript, hgnc_symbol) %>% summarise(mean_length = mean(Length))
summarise(data_transcript, mean_length = mean(Length))
group_by(data_transcript, hgnc_symbol) %>% summarise_if(is.numeric, funs(mean(.)))
```

# Combining tables with dplyr

## Combine tables
- Combine variables (i.e. columns)
    + `bind_cols()`
    + `left_join()`
    + `right_join()`
    + `inner_join()`
    + `full_join()`
    
- Combine cases (i.e. rows)
    + `bind_rows()`
    + `intersect()`
    + `setdiff()`
    + `union()`
  
## Combine columns

- `bind_cols()` to paste tables beside each other

- Syntax: 
```{}
bind_cols(data1, data2)
```
- Examples:   

```{}  
data1 <- select(data_transcript, hgnc_symbol, 
                  ensembl_gene_id, ensembl_transcript_id)
data2 <- select(data_transcript, -hgnc_symbol, -ensembl_gene_id)
glimpse(data1)
glimpse(data2)

#looks good!  

data3 <- bind_cols(data1, data2)
```
  
## Combine columns {.smaller}

- `left_join()` to join matching values from the right dataset to the left dataset

- Syntax: 

```{}
left_join(data1, data2)
left_join(data1, data2, by = ...)
left_join(data1, data2, by = c("col1" = "col2"))
```
- Examples:   

```{}  
data1 <- data1[c(1:8),] #dropping some rows
glimpse(data1)
data2 <- data2[c(3:15), ] #dropping some rows
glimpse(data2)

data3 <- left_join(data1, data2)
data4 <- right_join(data1, data2)
data5 <- inner_join(data1, data2)
data6 <- full_join(data1, data2)
```  

## Combine rows

- `bind_rows()` to paste tables below each other as they are. 

- More Examples:   

```{}  
data1 <- data_transcript[c(1:8), ] # subset of data_transcript
data2 <- data_transcript[c(8:18), ] # a different subset of data_transcript

data3 <- bind_rows(data1, data2)
```

## Combine rows

- `intersect()` to find rows that appear in both tables. 

- Syntax: 

```{}
intersect(data1, data2)
```
- Examples:   

```{}
data4 <- intersect(data1, data2)
data5 <- setdiff(data1, data2) # rows that appear in x, but not y
data6 <- setdiff(data2, data1) # flip by reversing the order of datasets
data7 <- union(data1, data2) # rows that appear in x or y
```

# tidyr

## tidyr verbs

- `spread`
- `gather`
- `separate`
- `unite`

## `spread` & `gather`

- When do you use it?
- When you want to change the **shape** of your data. 

## `spread`

- `spread()` takes two columns (key & value), and spreads into multiple columns: it makes “long” data wider
- Syntax: spread(data = ..., key = ..., value = ...)
- Example: 

## `gather`

- `gather()`  takes multiple columns, and gathers them into key-value pairs: it makes “wide” data longer
- Syntax: gather(data = ..., key = ..., value = ..., <columns-to-gather>)

## `separate` & `unite`

- When do you use it?
- When you want to split or combine the **contents** of a cell in your data. 

## `separate`

- `separate()` turns a single character column into multiple columns by splitting the values of the column wherever a separator character appears.
- Syntax: separate(data = ..., col = ..., into = ..., sep = ...)

## `unite`

- `unite()` combines multiple columns into a single column.
- Syntax: unite(data = ..., col = ..., <columns-to-unite>, sep = ...)

# Your turn {.biggest}

## Acknowledgements

Most images, especially the cheat sheets, are from www.rstudio.com