---
title: "Class 3b: ggplot2 revisited"
author: "Neelanjan Mukherjee"
date: "2/12/2019"
output: html_document
---
```{r , include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```




```{r setup,echo=FALSE, message=FALSE}


# We need to load all the libraries we think are needed for this exercise.
library(here) # setting wdir
library(tidyverse) # data-warngling
library(reshape2) # converting data from wide-to-long and long-to-wide
library(ggplot2) # plotting
library(ggthemes) # pretty plotting themes
library(scales) # nice annotating of x/y axes in plots
# install.packages("cowplot")
library(cowplot) # nice plots
```



```{r tidy data}

# set your working directory
setwd("~/Documents/molb7910-master/")

# read in dux data
dux_data <- read_csv(file ="3-class3/data/data_transcript.csv")

# remove the first column of `dux_data`
dux_data <- dux_data[,-1]

# let's see what is in dux data
names(dux_data)

# sum of counts for all transcrips of a given gene
dux_gene <- dux_data %>% select(hgnc_symbol, dux4_target, hour00_rep1:hour14_rep3) %>% group_by(hgnc_symbol,dux4_target) %>% summarise_all(funs(sum))


# let's log transform all counts data
# dux_log <- dux_gene %>%  mutate_at(.funs = funs(log = log10(. + 1)), .vars = vars(hour00_rep1:hour14_rep3)) %>% select(hgnc_symbol, dux4_target, hour00_rep1_log:hour14_rep3_log)


# ok now lets convert from wide-to-long
long_dux <- melt(dux_gene)
colnames(long_dux) <- c("gene","target","time","counts")

# i want a column for time and a separate one for replicate
long_dux_plot <- long_dux %>% separate(col = time, into = c("time","rep"))


```


## 1. Plotting data with ggplot2 

A popular approach to plotting in R is [ggplot2](https://ggplot2.tidyverse.org/). While graphs produced using ggplot2 defaults are (imo) not quite publication qualitym with a few tweaks you can pretty much get there! I find [STHDA](http://www.sthda.com/english/wiki/be-awesome-in-ggplot2-a-practical-guide-to-be-highly-effective-r-software-and-data-visualization) to be a great resource for `ggplot2` and many other `R` coding questions.

*** 
### A. The basic syntax of `ggplot()`  

`ggplot()`: build plots piece by piece
The concept of ggplot divides a plot into three different fundamental parts:

plot = __data__ + __aesthetics__ + __geometry__.  
<br/>
__data__: a data frame
<br/>
__aesthetics__: specify x and y variables, and other features - color, size, shape, etc.
<br/>
__geometry__: specify type of plots  - histogram, boxplot, line, density, dotplot, bar, etc. 
<br/>


### B. Producing different types of plots
    + Histograms and density plots
    + Scatter plots
    + Box-, violin-, and jitter-plots
    
### C. Additional considerations
    + Multiple plots
    + Themes
    + Titles, axes, and legends
    + Colors
    + Scales






### Exammple: making jitter plot step-by-step
We will make a variety of jitter plots now. Please see this [webpage](http://www.sthda.com/english/wiki/ggplot2-stripchart-jitter-quick-start-guide-r-software-and-data-visualization) for more information.

```{r jitter}
# we will use long_dux_plot, let's take a peak
head(long_dux_plot)

# basic plot
ggplot(data = long_dux_plot, aes(x = time, y = counts )) +  # `data` is 'long_dux_plot', `x` is 'time', `y` is 'counts'
  geom_jitter() # `geom` is 'geom_jitter()'

# tighten jitter
ggplot(data = long_dux_plot, aes(x = time, y = counts )) +  # `data` is 'long_dux_plot', `x` is 'time', `y` is 'counts'
  geom_jitter(width = 0.25) # `geom` is 'geom_jitter()'

# tighten jitter more - ahhh, thats better!
ggplot(data = long_dux_plot, aes(x = time, y = counts )) +  # `data` is 'long_dux_plot', `x` is 'time', `y` is 'counts'
  geom_jitter(width = 0.1) # `geom` is 'geom_jitter()'

# let's log that y-axis
ggplot(data = long_dux_plot, aes(x = time, y = counts)) +  # `data` is 'long_dux_plot', `x` is 'time', `y` is 'counts'
  geom_jitter(width = 0.1) + # `geom` is 'geom_jitter()'
  scale_y_log10() # make y-axis log10

# Neel has been so patient, let's make him happy and change the theme
ggplot(data = long_dux_plot, aes(x = time, y = counts)) +  # `data` is 'long_dux_plot', `x` is 'time', `y` is 'counts'
  geom_jitter(width = 0.1) + # `geom` is 'geom_jitter()'
  scale_y_log10() + # make y-axis log10
  theme_few() # my go-to theme

# sooo many themes
ggplot(data = long_dux_plot, aes(x = time, y = counts)) +  # `data` is 'long_dux_plot', `x` is 'time', `y` is 'counts'
  geom_jitter(width = 0.1) + # `geom` is 'geom_jitter()'
  scale_y_log10() + # make y-axis log10
  theme_fivethirtyeight() # hard not to like Nate Silver + Co.


# Neel has been so patient, let's make him happy and change the theme
ggplot(data = long_dux_plot, aes(x = time, y = counts)) +  # `data` is 'long_dux_plot', `x` is 'time', `y` is 'counts'
  geom_jitter(width = 0.1) + # `geom` is 'geom_jitter()'
  scale_y_log10() + # make y-axis log10
  theme_few() # my go-to theme

# let's be good scientists and add a title
ggplot(data = long_dux_plot, aes(x = time, y = counts)) +  # `data` is 'long_dux_plot', `x` is 'time', `y` is 'counts'
  geom_jitter(width = 0.1) + # `geom` is 'geom_jitter()'
  scale_y_log10() + # make y-axis log10
  theme_few() + # my go-to theme
  ggtitle("jitterplot  into my heart..") # title


# how to modify axis labels
ggplot(data = long_dux_plot, aes(x = time, y = counts)) +  # `data` is 'long_dux_plot', `x` is 'time', `y` is 'counts'
  geom_jitter(width = 0.1) + # `geom` is 'geom_jitter()'
  scale_y_log10() + # make y-axis log10
  theme_few() + # my go-to theme
  ggtitle("jitterplot  into my heart..") + # title
  ylab(label = "I love counts") + # my new y-axis name
  xlab(label = "Hear DUX4 Roar")  # my new x-axis name

# now let's color by the dux4 target
ggplot(data = long_dux_plot, aes(x = time, y = counts, color = target)) +  # `data` is 'long_dux_plot', `x` is 'time', `y` is 'counts', `color` by 'target'
  geom_jitter(width = 0.1) + # `geom` is 'geom_jitter()'
  scale_y_log10() + # make y-axis log10
  theme_few() + # my go-to theme
  ggtitle("jitterplot  into my heart..") + # title
  ylab(label = "I love counts") + # my new y-axis name
  xlab(label = "Hear DUX4 Roar")  # my new x-axis name

# not bad, but which genes are which!!
ggplot(data = long_dux_plot, aes(x = time, y = counts, color = target, shape = gene)) +  # `data` is 'long_dux_plot', `x` is 'time', `y` is 'counts', `color` by 'target', `shape` is 'gene'
  geom_jitter(width = 0.1) + # `geom` is 'geom_jitter()'
  scale_y_log10() + # make y-axis log10
  theme_few() + # my go-to theme
  ggtitle("jitterplot  into my heart..") + # title
  ylab(label = "I love counts") + # my new y-axis name
  xlab(label = "Hear DUX4 Roar")  # my new x-axis name

# ggplot was displeased because it doesn't have enough shapes!!
# let's swap target and gene
ggplot(data = long_dux_plot, aes(x = time, y = counts, color = gene, shape = target)) +  # `data` is 'long_dux_plot', `x` is 'time', `y` is 'counts', `color` by 'target', `shape` is 'gene'
  geom_jitter(width = 0.1) + # `geom` is 'geom_jitter()'
  scale_y_log10() + # make y-axis log10
  theme_few() + # my go-to theme
  ggtitle("jitterplot  into my heart..") + # title
  ylab(label = "I love counts") + # my new y-axis name
  xlab(label = "Hear DUX4 Roar")  # my new x-axis name



# this is annoying to look at
```


Now we will go more deeply into options with `ggplot2` that require creating a `ggplot2` object. We will also explore `facet_wrap()` which is a very useful tool.

```{r round two jitter}
# time for a new powerful trick called the `facet-wrap`
# but before we do this let's talk about making ggplot2 objects
# lets take the second to last plot we made and assign it to `p`

p_target <- ggplot(data = long_dux_plot, aes(x = time, y = counts, color = target)) +  # `data` is 'long_dux_plot', `x` is 'time', `y` is 'counts', `color` by 'target'
  geom_jitter(width = 0.1) + # `geom` is 'geom_jitter()'
  scale_y_log10() + # make y-axis log10
  theme_few() + # my go-to theme
  ggtitle("jitterplot  into my heart..") + # title
  ylab(label = "I love counts") + # my new y-axis name
  xlab(label = "Hear DUX4 Roar")  # my new x-axis name

p_target # prints plot

names(p_target) # what is in the object?

p_target$data # i can get the data back

p_target$mapping # my aesthetics

p_target$theme # that is way too much stuff!

p_target$labels # i see my labels


# remember 'p-target'
p_target

# let's say I want to make a separate `p_target`  plot for each gene
p_target + facet_wrap(gene ~ .) # `facet_wrap` gene 

# let's fix x-axis
p_target + facet_wrap(~gene) + # `facet_wrap` by gene 
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) # angle 60 degree

p_target + facet_wrap(~target) + # `facet_wrap` by target 
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) # angle 60 degree

p_target + facet_wrap(target~gene) + # `facet_wrap` target by gene
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) # angle 60 degree


p_gene <- ggplot(data = long_dux_plot, aes(x = time, y = counts, color = gene)) +  # `data` is 'long_dux_plot', `x` is 'time', `y` is 'counts', `color` by gene
  geom_jitter(width = 0.1) + # `geom` is 'geom_jitter()'
  scale_y_log10() + # make y-axis log10
  theme_few() + # my go-to theme
  ggtitle("jitterplot  into my heart..") + # title
  ylab(label = "I love counts") + # my new y-axis name
  xlab(label = "Hear DUX4 Roar")  # my new x-axis name

p_gene_target <- p_gene + facet_wrap(~target) + # `facet_wrap` target 
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) # angle 60 degree
p_target_gene <- p_target + facet_wrap(~gene) + # `facet_wrap` by gene 
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) # angle 60 degree


p_final <- plot_grid(p_gene_target, p_target_gene, labels = c("A","B"), nrow = 2)


save_plot(filename = "../images/final_plot.png", plot = p_final)
```



Finally, we will __always__ end our Rmarkdown documents with the sessing information:

```{r}
sessionInfo()
```


