---
title: "class4-homework.Rmd"
author: "Sujatha Jagannathan"
date: "February 28, 2019"
output: html_document
---
<br/>

The goal of this homework is to use `edgeR` to calculate fold change across two different conditions, combining `dplyr` verbs to transform the foldchange data, and then to plot it using `ggplot`.

We will also practice converting computational goals to "pseudocode" and then into "real code". This is an important aspect of coding. For this class, writing the pseudocode carries 50% of the points. So even if your real code does not work, if you think through how to accomplish the task at hand via pseudocode, that is worth half the points. 

Please execute the following tasks, knit the resulting .Rmd (`Lastname_class4-homework.Rmd`) and submit the .html output as your assignment.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE, eval = TRUE)
```

<br/>

#### 1. Load packages 
Packages: `tidyverse`, `skimr`, `here`, and `ggplot2` 
```{r, include = TRUE}
#load packages here

```

<br/>

#### 2. Set working directory
```{r}
#setwd/here

```

<br/>

#### 3. Import data
Path: "4-class4/data/input/data_gene-level.csv"
```{r}
#import data

```

<br/>

#### 4. Conduct edgeR analysis for the 8h sample, using the 0h sample as the control sample. 

Follow the same steps 1-10 we used for the 14h sample in our class. Note that the only changes you will make to this pipeline is to subset the geneExpr dataframe differently (step #1) and to save the final foldchange data with the appropriate name (step #10). However, please look up the steps in edgeR (use the references at the bottom of `class4-DGE.Rmd`) and try to understand them. 

```{r}
# edgeR analysis for 8h sample. 



```

<br/>

#### 5. Prepare the foldchange data for plotting. 

**Goal**:   

1. Move the rownames of `DUX4_14h$table` into a new column to enable the next steps. 

2. Combine the results with the original `data_gene_level` file using `ensembl_gene_id` to map the ids back to `hgnc_symbol`. We will need to gene symbols for subsequent plotting.


**Pseudocode**:  
1.   
2.   

**Actual code**: 
```{r}
# 1.   

# 2.   

```

<br/>

#### 6. Create the following plots to explore our data  

**Goal**:   

1. Create an M-A plot with the entire foldchange dataset. (M or log ratio here is`logFC`, and A or mean average here is `logCPM`). Color the points with `FDR > 0.05` in grey. 

2. Create a Volcano plot with the entire foldchange dataset. (Volcano plot is `logFC` vs. -log10(`FDR`). Color the points with `logFC > 2` blue and `logFC < 2` red. Draw a line at `FDR = 0.05`. 

3. Remake the volcano plot with the entire foldchange dataset, highlighting the following DUX4 target genes in one color and the nontargets in another color.
    - targets <- ZSCAN4, KHDC1L, RFPL1, DUXA, CCNA1, CLK1
    - nontargets <- RPS14, RPS3, RPL36, RPL28

**Pseudocode**:  
1.   
2.   
3.   

**Actual code**: 
```{r}
# plot 1. 

# plot 2. 

# plot 3. 

```

<br/>

#### 6. Print sessionInfo {.smallest}

```{r}
#sessionInfo

```
