---
title: "class4-homework.Rmd"
author: "Sujatha Jagannathan"
date: "February 28, 2019"
output: html_document
---
The goal of this homework is to get comfortable combining `dplyr` verbs to transform data and then to plot it using `ggplot`. 

Please execute the following tasks, knit the resulting .Rmd (Lastname_class4-homework.Rmd) and submit the .html output as your assignment.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE, eval = TRUE)
```

#### 1. Load packages 
Packages: `tidyverse`, `skimr`, `here`, and `ggplot2` 
```{r, include = TRUE}
#load packages here

```

#### 2. Set working directory
```{r}
#setwd/here

```

#### 3. Import data
Path: "4-class4/data/output/data_foldchange.csv"
```{r}
#import data

```

#### 4. select, group_by, summarise, mutate and filter

Goal: 

Pseudocode:

Actual code: 
```{r}
#select, group_by, summarise, mutate and filter

```

#### 5. ggplot

Goal:

Pseudocode:

Actual code: 
```{r}
# plot
```

#### 6. Print sessionInfo {.smallest}

```{r}
#sessionInfo

```
